---
title: "SciVal data prep with widgets"
author: "Tom Tullius"
date: "`r Sys.Date()`"
output: 
  html_notebook: 
    code_folding: hide
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

These code chunks process raw trainee data to produce a spreadsheet file (csv format) that can be used in the SciVal Shiny dashboard app.

Input to all code chunks is a comma-separated value (csv) format spreadsheet file.\


### Step 1: 
Combine a csv file containing trainee metadata with a csv file containing the trainee name and Scopus ID for each trainee. 

Associating a Scopus ID with each trainee allows the SciVal Shiny dashboard app to query the SciVal database and retrieve publication data for trainees.

The resulting file can be named and downloaded as a csv file to your computer for subsequent use.

```{r}

library (dplyr)
library (tidyverse)

##########. function to load a file from an interactive file input. #######


load_file <- function(path) {
  vroom::vroom(path, delim = ",")
}

######. end function      ######


```



```{r}

##########. interactive file selection. #########

fileInput("file1",
                "Choose CSV File with Trainee metadata",
                accept = ".csv"
      )

fileInput("file2",
                "Choose CSV File with Trainee names and Scopus ID's",
                accept = ".csv"
      )

DT::dataTableOutput("meta_table")


############.  load csv files into dataframes.  ###########

meta_df <- reactive({
  
  load_file(input$file1$datapath)
  
})


ID_df <- reactive({
  
  load_file(input$file2$datapath)
  
})


##########.  produce dataframe with Scopus ID's joined to trainee metadata.  #########

df12  <- reactive({
  
left_join(meta_df(), ID_df(), by = 'Author') |> select(-starts_with("Student"), -"first")

})

#df12

output$meta_table <- DT::renderDataTable(DT::datatable({
  
    df <- df12()
    df
    
  }))


#########. download csv file with trainee metadata joined to Scopus ID for each trainee.  ###########


downloadHandler(
    filename = function() {
      paste0(tools::file_path_sans_ext(input$file$name), ".csv")
    },
    content = function(file) {
      vroom::vroom_write(df12(), file, delim = ",")
    }
  )

```

### Step 2.


Produce a dataframe with each row containing a trainee name, Scopus ID, and number of papers for each year from 1998-2021.


The papers data come from a set of csv files downloaded from SciVal, each including data for up to 75 trainees.
```{r}


##########. interactive file selection. #########


fileInput("file3",
                "Choose CSV File with trainees 1 - 75 from SciVal",
                accept = ".csv"
      )

fileInput("file4",
                "Choose CSV File with trainees 76 - 150 from SciVal",
                accept = ".csv"
      )

fileInput("file5",
                "Choose CSV File with trainees 150 - 225 from SciVal",
                accept = ".csv"
      )


DT::dataTableOutput("table")


############.  load csv files into dataframes.  ###########

pubs1_df <- reactive({
  
  load_file(input$file3$datapath)
  
})


pubs2_df <- reactive({
  
  load_file(input$file4$datapath)
  
})

pubs3_df <- reactive({
  
  load_file(input$file5$datapath)
  
})

##########.  produce a dataframe with with each row containing trainee name, Scopus ID, and number of papers for each year from 1998-2021  #########

full_pubs_df  <- reactive({
  
  all_pubs_df <- union(pubs1_df(), pubs2_df()) |>  union(pubs3_df())
  full_pubs_df <- inner_join(df12(), all_pubs_df, by = 'ID')

})


output$table <- DT::renderDataTable(DT::datatable({
  
    df <- full_pubs_df()
    df
    
  }))
       

```

```{r}

#########. download csv file with each row containing trainee name, Scopus ID, and number of papers for each year from 1998-2021.  ###########


downloadHandler(
    filename = function() {
      paste0(tools::file_path_sans_ext(input$file$name), ".csv")
    },
    content = function(file) {
      vroom::vroom_write(full_pubs_df(), file, delim = ",")
    }
  )
```