---
title: "SciVal data prep with widgets"
author: "Tom Tullius"
date: "`r Sys.Date()`"
output: 
  html_notebook: 
    code_folding: hide
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

These code chunks process raw trainee data to produce a spreadsheet file (CSV format) that can be used in the SciVal Shiny dashboard app.

The purpose of this code is to join an in-house-maintained file that contains various metadata for a list of trainees, with a file exported from SciVal containing data on Scholarly Output (paper numbers) for the same list of trainees.

The two files are joined by trainee Scopus ID's, which must be included in each file.


### Step 1: 
Load a raw CSV file that was exported from SciVal with a set of trainee names, their Scopus ID's, and their Scholarly Output (numbers of papers over a range of years,  as specified in the SciVal search). 
Headers are automatically stripped off - no pre-processing of this file is needed.
Scholarly Output can be total papers, or first-author papers, or last/corresponding author papers, depending on the configuration of the SciVal search.

### Step 2: 
Load a CSV file containing trainee metadata. This file MUST include Scopus ID's for each trainee, to allow the Scholarly Output dataframe to be joined with the metadata dataframe. 
Metadata can include gender, URM status, year of PhD start, year of PhD end, months in training, type of first-year fellowship, mentor name, current employer, job title, zip code of job, etc.

The resulting file, which joins metadata with Scholarly Output, can be named and downloaded as a CSV file to your computer for subsequent use.

```{r}

library(plyr)
library (dplyr)
library (tidyverse)
library(janitor)
library(vroom)

##########. function to load a SciVal- produced file from an interactive file input. 
#######  strips out extraneous header lines and Tags column, rename id variable, 
#######  remove extra column at end, and separate last and first names into two variables


load_SciVal_file <- function(path) {
  vroom::vroom(path, delim = ",", skip = 13, col_select = -...3, na = c("-", "NA")) |> 
  dplyr::rename(id = ...2) |> 
  select(!last_col()) 
}

######. end function      ######

########## function to load a trainee metadata csv file (containing SciVal ID's) from an interactive file input #######


load_metadata_file <- function(path) {
  vroom::vroom(path, delim = ",", na = c("-", "NA")) |>
    clean_names()
}

######. end function      ######


```



```{r}

##########. interactive file selection. #########

fileInput("file1",
                "Choose CSV File with trainee metadata",
                accept = ".csv"
      )

fileInput("file2",
                "Choose SciVal CSV File with trainee names, Scopus ID's, and numbers of papers over a range of years",
                accept = ".csv"
      )

DT::dataTableOutput("meta_table")


############.  load csv files into dataframes.  ###########

meta_df <- reactive({
  
  req(input$file1)
  load_metadata_file(input$file1$datapath)
  
})


SciVal_df <- reactive({
  
  req(input$file2)
  load_SciVal_file(input$file2$datapath)
  
})

###############. combine metadata file with SciVal-produced file, by Scival ID

meta_id_pubs_df  <- reactive({
  
  meta_id_pubs_df <- inner_join(meta_df(), SciVal_df(), by = 'id') |>
    select(!...1) |>
    mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))
  

})

output$meta_table <- DT::renderDataTable(DT::datatable({
  
    meta_id_pubs_df_table <- meta_id_pubs_df()
    meta_id_pubs_df_table
    
  }))


#########. download csv file with trainee metadata joined to papers by year for each trainee.  ###########


downloadHandler(
    filename = function() {
      paste0(tools::file_path_sans_ext(input$file$name), ".csv")
    },
    content = function(file) {
      vroom::vroom_write(meta_id_pubs_df(), file, delim = ",")
    }
  )

```